\begin{center}
    \textbf{\huge{\uppercase{Engineering Physics}}}
    \\
    \vspace{.5cm}
    \textbf{\large{SH 402}}
\end{center}

\noindent\textbf{Lecture\ \ \ : 4} \hfill \textbf{Year : I } \\
\textbf{Tutorial \ : 1} \hfill \textbf{Part : I } \\
\textbf{Practical : 2}  \\

\par
\noindent 
\section*{Course Objective:}
To provide the concept and knowledge of physics with the emphasis of present day application.

\begin{enumerate}
    \item \textbf{Oscillation \hfill (7 hours)}
    \begin{enumerate}
        \item Mechanical Oscillation: Introduction
        \item Free oscillation
        \item Damped oscillation
        \item Forced mechanical oscillation
        \item EM Oscillation: Free, damped and forced electromagnetic oscillation
    \end{enumerate}
    
    \item \textbf{Wave Motion \hfill (3 hours)}
    \begin{enumerate}
        \item Waves and particles
        \item Progressive wave
        \item Energy, power and intensity of progressive wave
    \end{enumerate}
    
    \item \textbf{Acoustics \hfill (3 hours)}
    \begin{enumerate}
        \item Reverberation
        \item Sabine' Law
        \item Ultrasound and its applications
    \end{enumerate}
    
    \item \textbf{Physical Optics \hfill (12 hours)}
    \begin{enumerate}
        \item Interference
        \begin{enumerate}
            \item Intensity in double slit interference
            \item Interference in thin films
            \item Newton's rings
            \item Hadinger fringes
        \end{enumerate}
        
        \item Diffraction
        \begin{enumerate}
            \item Fresnel and Fraunhoffer's diffraction
            \item Intensity due to a single slit
            \item Diffraction grating
            \item X-ray diffraction, x-ray for material test
        \end{enumerate}
        
        \item Polarization
        \begin{enumerate}
            \item Double refraction
            \item Nichol prism, wave plates
            \item Optical activity, specific rotation
        \end{enumerate}
    \end{enumerate}
    
    \item \textbf{Geometrical Optics \hfill (3 hours)}
    \begin{enumerate}
        \item Lenses, combination of lenses
        \item Cardinal points
        \item Chromatic aberration
    \end{enumerate}
    
    \item \textbf{Laser and Fiber Optics \hfill (4 hours)}
    \begin{enumerate}
        \item Laser production
        \begin{enumerate}
            \item He-Ne laser
            \item User of laser
        \end{enumerate}
        
        \item Fiber Optics
        \begin{enumerate}
            \item Self focusing
            \item Applications of optical fiber
        \end{enumerate}
    \end{enumerate}
    
    \item \textbf{Electrostatics \hfill (8 hours)}
    \begin{enumerate}
        \item Electric charge and force
        \item Electric field and potential 
        \item Electrostatic potential energy
        \item Capacitors, capacitor with dielectric
        \item Charging and discharging of a capacitor
    \end{enumerate}
    
    \item \textbf{Electromagnetism \hfill (11 hours)}
    \begin{enumerate}
        \item Direct current: Electric current
        \begin{enumerate}
            \item Ohm's law, resistance and resistivity
            \item Semiconductor and superconductor
        \end{enumerate}
        
        \item Magnetic fields
        \begin{enumerate}
            \item Magnetic force and Torque
            \item Hall effect
            \item Cyclotron, synchrotron
            \item Biot-savart law
            \item Ampere's circuit law; magnetic fields straight conductors
            \item Faraday's laws, Induction and energy transformation, induced field
            \item LR circuit, induced magnetic field
            \item Displacement current
        \end{enumerate}
    \end{enumerate}
    
    \item \textbf{Electromagnetic waves \hfill (5 hours)}
    \begin{enumerate}
        \item Maxwell's equations
        \item Wave equations, speed
        \item E and B fields
        \item Continuity equation
        \item Energy transfer
    \end{enumerate}
    
    \item \textbf{Photon and matter waves \hfill (5 hours)}
    \begin{enumerate}
        \item Quantization of energy
        \item Electrons and matter waves
        \item Schrodinger wave equation
        \item Probability distribution
        \item One dimensional potential well
        \item Uncertainty principle
        \item Barrier tunneling
    \end{enumerate}
\end{enumerate}

\section*{Practical:}
\begin{enumerate}
    \item To determine the acceleration due to gravity and radius of gyration of the bar about an axis passing through its center of gravity.
    \item To determine the value of modulus of elasticity of the materials given, and moment of inertia of a circular disc using torsion pendulum.
    \item To determine the angle of prism and dispersive power of materials of the prism using spectrometer
    \item To determine the wavelength of sodium light by Newton's rings.
    \item To determine the wavelength of He-Ne laser light and use it to measure the thickness of a thin wire by diffraction of light.
    \item To study the variation of angle of rotation of plane of polarization using concentration of the cane sugar solution.
    \item To determine the specific rotation of the cane sugar solution using polarimeter.
    \item To determine the low resistance of a given wire by Carey Foster bridge and to determine the resistance per unit length of the wire of the bridge.
    \item To determine the capacitance of a given capacitor by charging and discharging through resistor.
    \item To plot a graph between current and frequency in a LRC series circuit and find the resonant frequency and quality factor.
    \item To determine dielectric constant of a given substance and study its variation with frequency by resonance method.
    \item To determine the susceptibility of a solution of given materials by Quinkes method.
    \item To study the electric field mapping.
    
\end{enumerate}


\section*{References:}
\begin{enumerate}
    \item Halliday, Resnick, Walker, ``Fundamentals of Physics", John Wiley \& Sons. Inc.
    \item Sapkota, Pokharel, Bhattarai, ``Fundamental of Engineering Physics", Benchmark Publication.
    \item Brij Lal and Subrahmanyam, ``A text book of Optics", S. Chand Publisher.
    \item A. S. Basudeva, ``Modern Engineering Physics", S. Chand Publisher.
    \item R. K. Gaur and S. L. Gupta, ``Engineering Physics", Dhanpat Publisher.
    \item Brij Lal and Subrahmanyam, ``Waves and Oscillation", S. Chand Publisher.
\end{enumerate}